This file will complete the first regression needed for the flow data: a limited difference-in-difference analysis of the Kyrgyz migrant flows (region-by-year 2021-23) over time.

Of interest:
- significance and magnitude of the coefficients on time fixed effect dummies --> quantify impact of war onset and continuation
- selection, significance, and magnitude of covariates --> understand pull factors for kyrgyz migrants to russia

```{r}
region_year_data <- read_csv("region-year-data.csv") %>%
  select(-contains("mediazona"), -contains("count_"), -contains("pct_"), -contains("ratio_")) %>%
#remove yandex trends & mediazona data
  select(-Outflow, -Netflow) %>%
  filter(!is.na(Inflow))  # Remove rows where Inflow is NA, if any
```

Prep by removing NAs
```{r}
library(dplyr)

# Identify rows with NA values and return the Year, Region, and the variable(s) with NA
na_data <- region_year_data %>%
  tidyr::gather(key = "variable", value = "value", -Year, -Region) %>%
  filter(is.na(value)) %>%
  select(Year, Region, variable)

# View the result
head(na_data)

#only lag_minwagevalue/lag_log_minwagevalue. Not that important, let's drop them. 
region_year_data <- region_year_data %>% 
  select(-lag_minwagevalue, -lag_log_minwagevalue)

# Identify rows with NA values and return the Year, Region, and the variable(s) with NA
na_data <- region_year_data %>%
  tidyr::gather(key = "variable", value = "value", -Year, -Region) %>%
  filter(is.na(value)) %>%
  select(Year, Region, variable)

# View the result
head(na_data) #no NAs left! 
```

Prep by removing time trends (sample too small) and adding a post-war dummy. 
```{r}
region_year_data <- region_year_data %>%
  # Remove the specified columns
  select(-c(t2, t3, exp_t, sin_t)) %>%
  # Add the postwar dummy variable
  mutate(postwar = if_else(Year >= 2022, 1, 0)) %>%
  mutate(t = t-1) #t should count up from 2022, not 2021. 
```

#DID to see changes over time

Run stepwise regression
```{r}
# Create the formula for the model (Inflow as the dependent variable)
formula <- as.formula(paste("Inflow ~", 
                            paste(names(region_year_data)[!(names(region_year_data) %in% c('Region', 'Year', 'Inflow'))], collapse = " + "),
                            "+ factor(Year) + factor(Region)"))

# Fit the full model using all covariates and fixed effects
full_model <- lm(formula, data = region_year_data)

# Perform stepwise regression using AIC for model selection (both directions)
stepwise_model <- step(full_model, direction = "both", trace = 1)

# View the summary of the selected model
summary(stepwise_model)
capture.output(summary(stepwise_model), file = "1_DID_inflow_analysis_summary.txt")
```

Jotting down some initial analysis:
t: 5.546e+02 *  --> indicates that as war continues there are more Kyrgyz migrants flowing into Russia, large effect (~500 migrants per year at regional level)
all population variables were very impactful. City pop was generally positive and rural pop generally negative.
a higher minimum wage value was associated with a lower Kyrgyz migrant inflow.
lagged average nominal gross wages also had a negative association --> lower wage areas 
postwar dummy not significant, but negative impact expected (~ -250 migrants, somewhat offsetting impact of t)
Not sig but a selected covariate: change_in_output (+), change_in_real_avg_wages (-), changeinindustrialproduction_mineral_extraction (-), lag_changeinindustrialproduction_water_utilities (-)

#Mediazona casualties (labor substitution analysis)
We simply add mediazona data back into the mix and see if it is selected and what its coefficients look like. 

```{r}
region_year_data <- read_csv("region-year-data.csv") %>%
  select(-contains("count_"), -contains("pct_"), -contains("ratio_")) %>%
#remove yandex trends but KEEP mediazona
  select(-Outflow, -Netflow) %>%
  filter(!is.na(Inflow))  # Remove rows where Inflow is NA, if any
```

Same as above: Prep by removing NAs
```{r}
library(dplyr)

# Identify rows with NA values and return the Year, Region, and the variable(s) with NA
na_data <- region_year_data %>%
  tidyr::gather(key = "variable", value = "value", -Year, -Region) %>%
  filter(is.na(value)) %>%
  select(Year, Region, variable)

# View the result
head(na_data)

#only lag_minwagevalue/lag_log_minwagevalue. Not that important, let's drop them. 
region_year_data <- region_year_data %>% 
  select(-lag_minwagevalue, -lag_log_minwagevalue)

# Identify rows with NA values and return the Year, Region, and the variable(s) with NA
na_data <- region_year_data %>%
  tidyr::gather(key = "variable", value = "value", -Year, -Region) %>%
  filter(is.na(value)) %>%
  select(Year, Region, variable)

# View the result
head(na_data) #no NAs left! 
```

Same as above: Prep by removing time trends (sample too small) and adding a post-war dummy. 
```{r}
region_year_data <- region_year_data %>%
  # Remove the specified columns
  select(-c(t2, t3, exp_t, sin_t)) %>%
  # Add the postwar dummy variable
  mutate(postwar = if_else(Year >= 2022, 1, 0)) %>%
  mutate(t = t-1) %>% #t should count up from 2022, not 2021. 
  mutate(interaction_logmediazona_t = log_mediazona*t)
```

Run stepwise regression, now including mediazona.
```{r}
# Create the formula for the model (Inflow as the dependent variable)
formula <- as.formula(paste("Inflow ~", 
                            paste(names(region_year_data)[!(names(region_year_data) %in% c('Region', 'Year', 'Inflow'))], collapse = " + "),
                            "+ factor(Year) + factor(Region)"))

# Fit the full model using all covariates and fixed effects
full_model <- lm(formula, data = region_year_data)

# Perform stepwise regression using AIC for model selection (both directions)
stepwise_model <- step(full_model, direction = "both", trace = 1)

# View the summary of the selected model
summary(stepwise_model)
capture.output(summary(stepwise_model), file = "2_mediazona_inflow_analysis_summary.txt")
```

Initial thoughts:
mediazona, log_mediazona, and lag_log_mediazona (small) all have significant negative impacts, but lag_mediazona and the interaction with t has a positive impact. These are all significant. 
This is evidence for a delayed substitution effect, since the lags are more significant, but the war overall driving migrants from Russia back to Kyrgyzstan.

