This file merges the data for the various predictors together to prep them for the elastic net.

Note: ChatGPT was used to generate some base code, but the code below is otherwise the author's own creation. 

Load in packages:
```{r packages}
library(readr)
library(readxl)
library(dplyr)
library(tidyr)
```

Load in data:
```{r}
Yandex_keywords_annual <- read_csv("~/Documents/thesis/Search queries data/Yandex Wordstat queries/Yandex_keywords_annual.csv")
```

Load in the KGZ migrant data (pulled from Rosstat):
```{r}
KGZ_migration_flows <- read_excel("~/Documents/thesis/Rosstat_KGZ_migration_flows.xlsx", skip = 2)
```

Clean the KGZ migrant data:
```{r}
#Handling the data header
colnames(KGZ_migration_flows)[1] <- "Region"
years <- as.character(KGZ_migration_flows[1, -1]) 
flow_types <- as.character(KGZ_migration_flows[2, -1])
KGZ_migration_flows <- KGZ_migration_flows[-c(1, 2), ]

#Long format data
KGZ_migration_flows <- KGZ_migration_flows %>%
  gather(key = "Year_Flow", value = "Value", -Region) %>%
  separate(Year_Flow, into = c("Year", "Flow_Type"), sep = "_") %>%
  spread(key = "Flow_Type", value = "Value")

#Check structure
head(KGZ_migration_flows)
```

Add transformations on Yandex keyword data:
```{r}
Yandex_keywords_annual <- Yandex_keywords_annual %>%
  filter(Year >= 2018) #because December 2017 was included — not comparable

#Dictionary object for future analysis (after adding more keywords)
dict <- list(
  "Work" = c("KGZ" = "Jumush", "RUS" = "Rabota", "ENG" = "Work"),
  "Work Contract" = c("KGZ" = "Emgek Kelishimi", "RUS" = "Trudovoi Dogovor", "ENG" = "Work Contract")
)

#Apply transformations on each variable
Yandex_keywords_annual <- Yandex_keywords_annual %>%
  # Log+1 Transformation
  mutate(across(starts_with("Jumush") | starts_with("Rabota"), ~ log(. + 1), .names = "log_{col}")) %>%
  
  #Ratio variable
  mutate(
    Work_ratio = ifelse(!is.na(Jumush) & !is.na(Rabota), Jumush / Rabota, NA),
    Work_diff = ifelse(!is.na(Rabota) & !is.na(Jumush), Rabota - Jumush, NA)
  ) %>%
  
  #Lag
  group_by(Region) %>%
  arrange(Region, Year) %>%
  mutate(
    Jumush_lag = lag(Jumush, 1),
    Rabota_lag = lag(Rabota, 1),
    
  #Percentage change
    Jumush_change_pct = ifelse(!is.na(Jumush_lag) & Jumush_lag != 0, (Jumush - Jumush_lag) / Jumush_lag * 100, 0),
    Rabota_change_pct = ifelse(!is.na(Rabota_lag) & Rabota_lag != 0, (Rabota - Rabota_lag) / Rabota_lag * 100, 0),
    Work_ratio_pct = ifelse(!is.na(lag(Work_ratio)) & lag(Work_ratio) != 0, (Work_ratio - lag(Work_ratio)) / Work_ratio * 100, 0)
  ) %>%
  ungroup()

#Check structure
head(Yandex_keywords_annual)
```

Checking region spellings for standardization, and creating a dictionary:
```{r}
#head(unique(KGZ_migration_flows$Region))
#head(unique(Yandex_keywords_annual$Region))

region_dict <- data.frame(
  Russian = c(
    "Алтайский край", "Амурская область", "Архангельская область", 
    "Архангельская область                       без автономного округа", 
    "Астраханская область", "Белгородская область", "Брянская область", 
    "Владимирская область", "Волгоградская область", "Вологодская область", 
    "Воронежская область", "г.Москва", "г.Санкт-Петербург", 
    "г.Севастополь", "Дальневосточный федеральный округ", 
    "Еврейская автономная область", "Забайкальский край", 
    "Ивановская область", "Иркутская область", 
    "Кабардино-Балкарская Республика", "Калининградская область", 
    "Калужская область", "Камчатский край", 
    "Карачаево-Черкесская Республика", "Кемеровская область", 
    "Кировская область", "Костромская область", "Краснодарский край", 
    "Красноярский край", "Курганская область", "Курская область", 
    "Ленинградская область", "Липецкая область", "Магаданская область", 
    "Московская область", "Мурманская область", "Ненецкий автономный округ", 
    "Нижегородская область", "Новгородская область", 
    "Новосибирская область", "Омская область", "Оренбургская область", 
    "Орловская область", "Пензенская область", "Пермский край", 
    "Приволжский федеральный округ", "Приморский край", 
    "Псковская область", "Республика Адыгея", "Республика Алтай", 
    "Республика Башкортостан", "Республика Бурятия", 
    "Республика Дагестан", "Республика Ингушетия", "Республика Калмыкия", 
    "Республика Карелия", "Республика Коми", "Республика Крым", 
    "Республика Марий Эл", "Республика Мордовия", "Республика Саха (Якутия)", 
    "Республика Северная Осетия-Алания", "Республика Татарстан", 
    "Республика Тыва", "Республика Хакасия", "Ростовская область", 
    "Рязанская область", "Самарская область", "Саратовская область", 
    "Сахалинская область", "Свердловская область", 
    "Северо-Западный федеральный округ", "Северо-Кавказский федеральный округ", 
    "Сибирский федеральный округ", "Смоленская область", "Ставропольский край", 
    "Тамбовская область", "Тверская область", "Томская область", 
    "Тульская область", "Тюменская область", 
    "Тюменская область                             без автономных округов", 
    "Удмуртская Республика", "Ульяновская область", 
    "Уральский федеральный округ", "Хабаровский край", 
    "Ханты-Мансийский автономный                 округ-Югра", 
    "Центральный федеральный округ", "Челябинская область", 
    "Чеченская Республика", "Чувашская Республика", 
    "Чукотский автономный округ", "Южный федеральный округ", 
    "Ямало-Ненецкий автономный округ", "Ярославская область"
  ),
  English = c(
    "Altai Krai", "Amur Oblast", "Arkhangel'skaya Oblast'", 
    "Arkhangelskaya Oblast without Autonomous Okrug", 
    "Astrakhan Oblast", "Belgorod Oblast", "Bryansk Oblast", 
    "Vladimir Oblast", "Volgograd Oblast", "Vologda Oblast", 
    "Voronezh Oblast", "Moscow", "Saint Petersburg", 
    "Sevastopol", "Far Eastern Federal District", 
    "Jewish Autonomous Oblast", "Zabaykalsky Krai", 
    "Ivanovo Oblast", "Irkutsk Oblast", 
    "The Kabardino-Balkar Republic", "Kaliningrad Oblast", 
    "Kaluga Oblast", "Kamchatka Krai", 
    "Karachay-Cherkess Republic", "Kemerovo Oblast", 
    "Kirov Oblast", "Kostroma Oblast", "Krasnodar Krai", 
    "Krasnoyarsk Krai", "Kurgan Oblast", "Kursk Oblast", 
    "Leningrad Oblast", "Lipetsk Oblast", "Magadan District", 
    "Moscow Oblast", "Murmansk Oblast", "Nenets Autonomous Okrug", 
    "Nizhny Novgorod Oblast'", "Novgorod Oblast", 
    "Novosibirsk Oblast", "Omsk Oblast", "Orenburg Oblast", 
    "Oryol Oblast", "Penza Oblast", "Perm Krai", 
    "Volga Federal District", "Primorsky Krai", 
    "Pskov Oblast", "Republic of Adygea", "Altai Republic", 
    "Republic of Bashkortostan", "Republic of Buryatia", 
    "Republic of Dagestan", "Republic of Ingushetia", "Republic of Kalmykia", 
    "Republic of Karelia", "Komi Republic", "Republic of Crimea", 
    "Mari El Republic", "Republic of Mordovia", "Sakha (Yakutia) Republic", 
    "Republic of North Ossetia — Alania", "Republic of Tatarstan", 
    "Tyva Republic", "Republic of Khakassia", "Rostov Oblast", 
    "Ryazan Oblast", "Samara Oblast", "Saratov Oblast", 
    "Sakhalin District", "Sverdlovsk Oblast", 
    "North-Western Federal District", "North Caucasian Federal District", 
    "Siberian Federal District", "Smolensk Oblast", "Stavropol Krai", 
    "Tambov Oblast", "Tver Oblast", "Tomsk District", 
    "Tula Oblast", "Tyumen Oblast", 
    "Tyumen Oblast without Autonomous Okrugs", 
    "Udmurt Republic", "Ulyanovsk Oblast", 
    "Ural Federal District", "Khabarovsk Krai", 
    "Khanty-Mansi Autonomous Okrug — Yugra", 
    "Central Federal District", "Chelyabinsk Oblast", 
    "Chechen Republic", "Chuvash Republic", 
    "Chukotka Autonomous Okrug", "Southern Federal District", 
    "Yamalo-Nenets Autonomous Okrug", "Yaroslavl Oblast"
  )
)
```

Merge datasets by region and year:
```{r}
#Replace Russian region names in KGZ_migration_flows with English names
KGZ_migration_flows <- KGZ_migration_flows %>%
  left_join(region_dict, by = c("Region" = "Russian")) %>%
  mutate(Region = English) %>%
  select(-English) %>%
  mutate(Year = as.numeric(Year)) 

#Merge by region and year
merged_data <- Yandex_keywords_annual %>%
  left_join(KGZ_migration_flows, by = c("Region", "Year"))

#Check structure
head(merged_data)
```

Add time trends:
```{r}
merged_data <- merged_data %>%
  mutate(
    t = Year - min(Year) + 1,  # Normalize t to start from 1
    t2 = t^2,
    t3 = t^3,
    log_t = log(t),
    exp_t = exp(t),
    sin_t = sin(t)
  )
```

Export:
```{r}
write.csv(merged_data, "/Users/michelle/Documents/thesis/merged_data.csv", row.names = FALSE)
```

