This file will produce summary tables for inclusion in the exploratory data analysis section.

Load in libraries.
```{r libraries}
library(readr)
library(dplyr)
library(knitr)
library(kableExtra)
library(webshot2) 
library(magrittr)
library(tidyr)

setwd("/Users/michelle/Documents/thesis")
```

Load in data (region-year). 
```{r}
ry_merged_data <- read_csv("region-year-data.csv") %>%
  filter(Year >= 2021,
         Region != "Russian Federation")

# Coerce all columns except Region and Year to numeric
ry_merged_data <- ry_merged_data %>%
  mutate(across(.cols = !c(Region, Year), .fns = as.numeric))
```

Summary statistics (region-year).
```{r}
summary_table <- ry_merged_data %>%
  select(-Region, -Year) %>%
  summarise(across(everything(), list(
    Mean = ~mean(., na.rm = TRUE),
    SD = ~sd(., na.rm = TRUE),
    Min = ~min(., na.rm = TRUE),
    Max = ~max(., na.rm = TRUE),
    N = ~sum(!is.na(.))
  ), .names = "{.col}_{.fn}")) %>%
  pivot_longer(cols = everything(),
               names_to = c("Variable", ".value"),
               names_pattern = "^(.*)_(Mean|SD|Min|Max|N)$")

#Table
kable_out <- summary_table %>%
  kable("html", caption = "Summary Statistics for Region-by-Year variables, 2021-24", digits = 2) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                full_width = F, 
                position = "center") %>%
  column_spec(1, bold = T) %>%
  add_header_above(c(" " = 1, "Statistics" = 5))

save_kable(kable_out, "summary_statistics.html")
```

Correlation matrix (region-year)
```{r}
# Load necessary packages
library(urca)
library(dplyr)

# Prepare data: Remove Region and Year columns
data_for_cointegration <- ry_merged_data %>%
  select(-Region, -Year)

# Drop rows with missing values
data_for_cointegration <- data_for_cointegration %>% drop_na()

cor_matrix <- cor(data_for_cointegration, use = "complete.obs")
print(cor_matrix)

# Save the kable as an HTML file
save_kable(kable(cor_matrix, caption = "Correlation Matrix of Variables", format = "html"), "correlation_matrix.html")
```

Find perfectly correlated pairs
```{r}
# Find the pairs of variables with perfect correlation (1 or -1)
perfectly_correlated <- which(abs(cor_matrix) == 1, arr.ind = TRUE)

# Remove self-correlation (diagonal elements)
perfectly_correlated <- perfectly_correlated[perfectly_correlated[, 1] != perfectly_correlated[, 2], ]

# Extract the variable names that are perfectly correlated
perfectly_correlated_vars <- apply(perfectly_correlated, 1, function(x) colnames(cor_matrix)[x])

# Print the pairs of variables with perfect correlation
perfectly_correlated_vars
```

Seems like nothing is perfectly correlated except for with itself and the absolute count variables between sum/average aggregation methods. 
This shouldn't be an issue because the methods will simply select one or none, but we'll keep in mind that we can drop the average variable if it causes issues. 
