Goal: merge manually-downloaded Yandex data into a single dataframe for each keyword.

```{r}
# Load necessary libraries
library(readr)
library(dplyr)
library(stringr)
library(lubridate)
library(tidyr)
```

Added some data in Russian and English, so we need a function to handle both:
```{r}
# Updated Function to parse date in both English and Russian formats
parse_period <- function(period) {
  # Check for Excel serial number format
  if (grepl("^\\d{5}$", period)) {
    date_serial <- as.Date(as.numeric(period), origin = "1899-12-30")
    return(date_serial)
  }

  # Try parsing English month-year format
  date_eng <- suppressWarnings(lubridate::my(period))
  if (!is.na(date_eng)) return(date_eng)

  # Try parsing Russian month-year format
  months_ru <- c("январь" = "January", "февраль" = "February", "март" = "March", 
                 "апрель" = "April", "май" = "May", "июнь" = "June", 
                 "июль" = "July", "август" = "August", "сентябрь" = "September", 
                 "октябрь" = "October", "ноябрь" = "November", "декабрь" = "December")

  # Replace Russian month names with English equivalents
  period_ru <- tolower(period)
  for (ru_month in names(months_ru)) {
    period_ru <- gsub(ru_month, months_ru[[ru_month]], period_ru)
  }
  date_ru <- suppressWarnings(lubridate::my(period_ru))
  if (!is.na(date_ru)) return(date_ru)

  # Handle Russian numeric date format (e.g., 31.12.2017)
  period_clean <- stringr::str_extract(period, "\\d{2}\\.\\d{2}\\.\\d{4}")
  date_ru_num <- suppressWarnings(lubridate::dmy(period_clean))
  if (!is.na(date_ru_num)) return(date_ru_num)

  # If all parsing fails, return NA
  return(NA)
}
```

Dictionary for translating Russian region names to latinized versions:
```{r}
# Create a named vector to map Russian region names to English equivalents
russian_to_english <- c(
  "Калужская область" = "Kaluga Oblast",
  "Костромская область" = "Kostroma Oblast",
  "Курская область" = "Kursk Oblast",
  "Липецкая область" = "Lipetsk Oblast",
  "Орловская область" = "Oryol Oblast",
  "Рязанская область" = "Ryazan Oblast",
  "Смоленская область" = "Smolensk Oblast",
  "Тамбовская область" = "Tambov Oblast",
  "Тверская область" = "Tver Oblast",
  "Тульская область" = "Tula Oblast",
  "Ярославская область" = "Yaroslavl Oblast",
  "Санкт-Петербург" = "Saint Petersburg",
  "Ленинградская область" = "Leningrad Oblast",
  "Архангельская область" = "Arkhangel'skaya Oblast'",
  "Вологодская область" = "Vologda Oblast",
  "Калининградская область" = "Kaliningrad Oblast",
  "Мурманская область" = "Murmansk Oblast",
  "Новгородская область" = "Novgorod Oblast",
  "Псковская область" = "Pskov Oblast",
  "Республика Карелия" = "Republic of Karelia",
  "Москва" = "Moscow",
  "Республика Коми" = "Komi Republic",
  "Кировская область" = "Kirov Oblast",
  "Нижегородская область" = "Nizhny Novgorod Oblast'",
  "Оренбургская область" = "Orenburg Oblast",
  "Пензенская область" = "Penza Oblast",
  "Пермский край" = "Perm Krai",
  "Республика Башкортостан" = "Republic of Bashkortostan",
  "Республика Марий Эл" = "Mari El Republic",
  "Республика Мордовия" = "Republic of Mordovia",
  "Республика Татарстан" = "Republic of Tatarstan",
  "Московская область" = "Moscow Oblast",
  "Самарская область" = "Samara Oblast",
  "Саратовская область" = "Saratov Oblast",
  "Удмуртская Республика" = "Udmurt Republic",
  "Ульяновская область" = "Ulyanovsk Oblast",
  "Чувашская Республика" = "Chuvash Republic",
  "Астраханская область" = "Astrakhan Oblast",
  "Волгоградская область" = "Volgograd Oblast",
  "Краснодарский край" = "Krasnodar Krai",
  "Республика Адыгея" = "Republic of Adygea",
  "Республика Калмыкия" = "Republic of Kalmykia",
  "Белгородская область" = "Belgorod Oblast",
  "Ростовская область" = "Rostov Oblast",
  "Алтайский край" = "Altai Krai",
  "Иркутская область" = "Irkutsk Oblast",
  "Кемеровская область (Кузбасс)" = "Kemerovo Oblast",
  "Красноярский край" = "Krasnoyarsk Krai",
  "Новосибирская область" = "Novosibirsk Oblast",
  "Омская область" = "Omsk Oblast",
  "Республика Алтай" = "Altai Republic",
  "Республика Тыва" = "Tyva Republic",
  "Республика Хакасия" = "Republic of Khakassia",
  "Брянская область" = "Bryansk Oblast",
  "Томская область" = "Tomsk District",
  "Амурская область" = "Amur Oblast",
  "Еврейская автономная область" = "Jewish Autonomous Oblast",
  "Забайкальский край" = "Zabaykalsky Krai",
  "Камчатский край" = "Kamchatka Krai",
  "Магаданская область" = "Magadan District",
  "Приморский край" = "Primorsky Krai",
  "Республика Бурятия" = "Republic of Buryatia",
  "Республика Саха (Якутия)" = "Sakha (Yakutia) Republic",
  "Сахалинская область" = "Sakhalin District",
  "Владимирская область" = "Vladimir Oblast",
  "Хабаровский край" = "Khabarovsk Krai",
  "Кабардино-Балкарская Республика" = "The Kabardino-Balkar Republic",
  "Карачаево-Черкесская Республика" = "Karachay-Cherkess Republic",
  "Республика Дагестан" = "Republic of Dagestan",
  "Республика Ингушетия" = "Republic of Ingushetia",
  "Республика Северная Осетия — Алания" = "Republic of North Ossetia — Alania",
  "Ставропольский край" = "Stavropol Krai",
  "Чеченская Республика" = "Chechen Republic",
  "Курганская область" = "Kurgan Oblast",
  "Свердловская область" = "Sverdlovsk Oblast",
  "Воронежская область" = "Voronezh Oblast",
  "Тюменская область" = "Tyumen Oblast",
  "Ханты-Мансийский автономный округ - Югра" = "Khanty-Mansi Autonomous Okrug — Yugra",
  "Челябинская область" = "Chelyabinsk Oblast",
  "Ямало-Ненецкий автономный округ" = "Yamalo-Nenets Autonomous Okrug",
  "Ивановская область" = "Ivanovo Oblast"
)
```

##Import the absolute counts

Convert the Yandex Wordstat data to a dataframe for each keyword. (takes 1-2 mins) 
```{r}
# List all folders (keywords)
folders <- list.dirs(full.names = TRUE, recursive = FALSE)

# Define the output directory
output_dir <- "/Users/michelle/Documents/thesis/Search queries data/Yandex Wordstat queries/clean keyword data"

# Ensure the directory exists
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}

# Loop through each folder
for (folder in folders) {
  setwd("/Users/michelle/Documents/thesis/Search queries data/Yandex Wordstat queries/raw keyword data")
  
  keyword <- basename(folder)  # Extract the folder name
  files <- list.files(folder, pattern = "*.csv", full.names = TRUE)  # List all CSV files
  
  keyword_df <- NULL  # Initialize an empty dataframe for the keyword
  
  for (file in files) {
    # Read the file
    temp_df <- read_delim(file, delim = ";", escape_double = FALSE, trim_ws = TRUE)
    
    # Extract the region title from column 4
    col4_name <- colnames(temp_df)[4] 
    
    # Regex patterns for English and Russian
    region_pattern_eng <- ".*\\d{2}\\.\\d{2}\\.\\d{4}, ([^,]+), all devices.*"
    region_pattern_ru <- ".*\\d{2}\\.\\d{2}\\.\\d{4}, ([^,]+), все устройства.*"
    
    region <- ifelse(grepl(region_pattern_eng, col4_name),
                     sub(region_pattern_eng, "\\1", col4_name),
                     sub(region_pattern_ru, "\\1", col4_name))
    
    # Handle special regions
    if (grepl("Moscow and Moscow Oblast, except Moscow", col4_name) || grepl("Москва и область, кроме Москва", col4_name)) {
      region <- "Moscow Oblast"
    } else if (grepl("Saint-Petersburg and Leningrad Oblast, except Saint Petersburg", col4_name) || grepl("Санкт-Петербург и Ленинградская область, кроме Санкт-Петербург", col4_name)) {
      region <- "Leningrad Oblast"
    }
    
    # Select relevant columns and rename them
    temp_df <- temp_df[, 1:2]  # Keep only "Period" and "Number of queries"
    temp_df[[2]] <- as.numeric(str_replace_all(temp_df[[2]], " ", ""))  # Remove spaces and convert to numeric
    colnames(temp_df) <- c("Period", region)  # Rename second column to the region
    
    # Merge into main dataframe
    if (is.null(keyword_df)) {
      keyword_df <- temp_df  # First file becomes the dataframe
    } else {
      keyword_df <- full_join(keyword_df, temp_df, by = "Period")  # Merge by "Period"
    }
  }
  
  # Convert "Period" to date format (supports English, Russian, and serial)
  keyword_df <- keyword_df %>%
    mutate(Period = sapply(Period, parse_period)) %>%
    mutate(across(-Period, ~ as.numeric(str_replace_all(., " ", "")))) 
    
    # Function to translate column names
  translate_columns <- function(df, mapping) {
    colnames(df) <- sapply(colnames(df), function(col) ifelse(col %in% names(mapping), mapping[col], col))
    return(df)
  }
  
  # Apply translation to the entire dataframe
  keyword_df <- translate_columns(keyword_df, mapping = russian_to_english)

  # Assign the dataframe to a variable named after the keyword
  assign(keyword, keyword_df, envir = .GlobalEnv)

  # Save the cleaned dataframe as a CSV
  write_csv(keyword_df, file.path(output_dir, paste0("count_", keyword, ".csv")))
}
```

##Import the percentages (same thing but with a different column of the raw data)

Convert the Yandex Wordstat data to a dataframe for each keyword. (takes 1-2 mins) 
```{r}
# List all folders (keywords)
folders <- list.dirs(full.names = TRUE, recursive = FALSE)

# Define the output directory
output_dir <- "/Users/michelle/Documents/thesis/Search queries data/Yandex Wordstat queries/clean keyword data"

# Ensure the directory exists
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}

# Loop through each folder
for (folder in folders) {
  setwd("/Users/michelle/Documents/thesis/Search queries data/Yandex Wordstat queries/raw keyword data")
  
  keyword <- basename(folder)  # Extract the folder name
  files <- list.files(folder, pattern = "*.csv", full.names = TRUE)  # List all CSV files
  
  keyword_df <- NULL  # Initialize an empty dataframe for the keyword
  
  for (file in files) {
    # Read the file
    temp_df <- read_delim(file, delim = ";", escape_double = FALSE, trim_ws = TRUE)
    
    # Extract the region title from column 4
    col4_name <- colnames(temp_df)[4] 
    
    # Regex patterns for English and Russian
    region_pattern_eng <- ".*\\d{2}\\.\\d{2}\\.\\d{4}, ([^,]+), all devices.*"
    region_pattern_ru <- ".*\\d{2}\\.\\d{2}\\.\\d{4}, ([^,]+), все устройства.*"
    
    region <- ifelse(grepl(region_pattern_eng, col4_name),
                     sub(region_pattern_eng, "\\1", col4_name),
                     sub(region_pattern_ru, "\\1", col4_name))
    
    # Handle special regions
    if (grepl("Moscow and Moscow Oblast, except Moscow", col4_name) || grepl("Москва и область, кроме Москва", col4_name)) {
      region <- "Moscow Oblast"
    } else if (grepl("Saint-Petersburg and Leningrad Oblast, except Saint Petersburg", col4_name) || grepl("Санкт-Петербург и Ленинградская область, кроме Санкт-Петербург", col4_name)) {
      region <- "Leningrad Oblast"
    }
    
    temp_df <- temp_df[, c(1, 3)]  # Keep "Period", "Number of queries", and "Percentage change"
    # Remove semicolons and spaces, then convert to numeric
temp_df[[2]] <- str_replace_all(temp_df[[2]], ";", "")  # Remove semicolons
temp_df[[2]] <- str_replace_all(temp_df[[2]], ",", ".")  # Replace commas with dots
temp_df[[2]] <- as.numeric(temp_df[[2]])  # Convert to numeric
    colnames(temp_df) <- c("Period", region, paste(region, "Percentage Change", sep = "_"))  # Rename columns
    
    # Merge into main dataframe
    if (is.null(keyword_df)) {
      keyword_df <- temp_df  # First file becomes the dataframe
    } else {
      keyword_df <- full_join(keyword_df, temp_df, by = "Period")  # Merge by "Period"
    }
  }
  
  # Convert "Period" to date format (supports English, Russian, and serial)
  keyword_df <- keyword_df %>%
    mutate(Period = sapply(Period, parse_period)) %>%
    mutate(across(-Period, ~ as.numeric(str_replace_all(., " ", "")))) 
    
    # Function to translate column names
  translate_columns <- function(df, mapping) {
    colnames(df) <- sapply(colnames(df), function(col) ifelse(col %in% names(mapping), mapping[col], col))
    return(df)
  }
  
  # Apply translation to the entire dataframe
  keyword_df <- translate_columns(keyword_df, mapping = russian_to_english)

  # Assign the dataframe to a variable named after the keyword
  assign(keyword, keyword_df, envir = .GlobalEnv)

  # Save the cleaned dataframe as a CSV
  write_csv(keyword_df, file.path(output_dir, paste0("pct_", keyword, ".csv")))
}
```

#Merge together (treating like a list of n*2 variables)

Now process and merge the keyword data together:
```{r}
# Set the directory for the cleaned keyword data
clean_data_dir <- "/Users/michelle/Documents/thesis/Search queries data/Yandex Wordstat queries/clean keyword data"

# List all the CSV files in the directory
files <- list.files(clean_data_dir, pattern = "*.csv", full.names = TRUE)

# Initialize an empty list to store the dataframes
long_data_list <- list()

# Loop through each file
for (file in files) {
  
  # Read the current CSV file
  temp_df <- read_csv(file)
  
  # Extract the keyword from the file name (without extension)
  keyword <- tools::file_path_sans_ext(basename(file))
  
  # Convert to long format
  temp_long_df <- temp_df %>%
    pivot_longer(cols = -Period, names_to = "region", values_to = keyword) %>%
    rename(Period = Period)  # Rename the 'Period' column explicitly
  
  # Store the long format data in the list
  long_data_list[[keyword]] <- temp_long_df
}

# Merge all datasets by 'Period' and 'region' (with outer join to keep all observations)
yandex_keywords_monthly <- Reduce(function(x, y) {
  merge(x, y, by = c("Period", "region"), all = TRUE)
}, long_data_list)

# View the result
head(yandex_keywords_monthly)
```

Annualize:
```{r}
yandex_keywords_monthly <- yandex_keywords_monthly %>%
  mutate(Period = Period - 17501) %>%
  mutate(Period = as_date(Period, origin = "2017-12-01")) %>%
  mutate(Period = format(Period, "%d-%m-%Y"))

# Extract the year from the 'Period' column and create a new 'Year' column
yandex_keywords_monthly <- yandex_keywords_monthly %>%
  mutate(Period = dmy(Period)) %>%  # Convert 'Period' to Date (Day-Month-Year)
  mutate(Year = year(Period))  # Extract the year

# Group by region and year, then sum and average the numeric data for each region/year
yandex_keywords_annual <- yandex_keywords_monthly %>%
  group_by(region, Year) %>%
  summarise(
    across(where(is.numeric), list(sum = ~sum(. , na.rm = TRUE), 
                                   average = ~mean(. , na.rm = TRUE))),
    .groups = "drop"
  )

# View the result
head(yandex_keywords_annual)
```

Drop percentage sums because they don't make sense:
```{r}
yandex_keywords_annual <- yandex_keywords_annual %>%
  select(-matches("pct.*sum"))  # Drop columns that have both 'pct' and 'sum' in their name
```

Export:
```{r}
write_csv(yandex_keywords_annual, file.path("/Users/michelle/Documents/thesis/Search queries data/Yandex Wordstat queries", "Yandex_keywords_annual.csv"))
```

