This file will merge together the region-year and region-month data by extracting it from several data files. 

Used ChatGPT to speed the process of managing so many different data formats. 

Load in packages
```{r}
library(readxl)
library(dplyr)
library(tidyr)
library(janitor)
library(zoo)
library(stringr)
library(lubridate)
library(tidyverse)

setwd("/Users/michelle/Documents/thesis")
```

##Set up region dictionary (standardizing names)
```{r}
region_dict <- data.frame(
  Russian = c(
    "Алтайский край", "Амурская область", "Архангельская область", 
    "Архангельская область                       без автономного округа", 
    "Астраханская область", "Белгородская область", "Брянская область", 
    "Владимирская область", "Волгоградская область", "Вологодская область", 
    "Воронежская область", "г.Москва", "г.Санкт-Петербург", 
    "г.Севастополь", "Дальневосточный федеральный округ", 
    "Еврейская автономная область", "Забайкальский край", 
    "Ивановская область", "Иркутская область", 
    "Кабардино-Балкарская Республика", "Калининградская область", 
    "Калужская область", "Камчатский край", 
    "Карачаево-Черкесская Республика", "Кемеровская область", 
    "Кировская область", "Костромская область", "Краснодарский край", 
    "Красноярский край", "Курганская область", "Курская область", 
    "Ленинградская область", "Липецкая область", "Магаданская область", 
    "Московская область", "Мурманская область", "Ненецкий автономный округ", 
    "Нижегородская область", "Новгородская область", 
    "Новосибирская область", "Омская область", "Оренбургская область", 
    "Орловская область", "Пензенская область", "Пермский край", 
    "Приволжский федеральный округ", "Приморский край", 
    "Псковская область", "Республика Адыгея", "Республика Алтай", 
    "Республика Башкортостан", "Республика Бурятия", 
    "Республика Дагестан", "Республика Ингушетия", "Республика Калмыкия", 
    "Республика Карелия", "Республика Коми", "Республика Крым", 
    "Республика Марий Эл", "Республика Мордовия", "Республика Саха (Якутия)", 
    "Республика Северная Осетия-Алания", "Республика Татарстан", 
    "Республика Тыва", "Республика Хакасия", "Ростовская область", 
    "Рязанская область", "Самарская область", "Саратовская область", 
    "Сахалинская область", "Свердловская область", 
    "Северо-Западный федеральный округ", "Северо-Кавказский федеральный округ", 
    "Сибирский федеральный округ", "Смоленская область", "Ставропольский край", 
    "Тамбовская область", "Тверская область", "Томская область", 
    "Тульская область", "Тюменская область", 
    "Тюменская область                             без автономных округов", 
    "Удмуртская Республика", "Ульяновская область", 
    "Уральский федеральный округ", "Хабаровский край", 
    "Ханты-Мансийский автономный                 округ-Югра", 
    "Центральный федеральный округ", "Челябинская область", 
    "Чеченская Республика", "Чувашская Республика", 
    "Чукотский автономный округ", "Южный федеральный округ", 
    "Ямало-Ненецкий автономный округ", "Ярославская область"
  ),
  English = c(
    "Altai Krai", "Amur Oblast", "Arkhangel'skaya Oblast'", 
    "Arkhangelskaya Oblast without Autonomous Okrug", 
    "Astrakhan Oblast", "Belgorod Oblast", "Bryansk Oblast", 
    "Vladimir Oblast", "Volgograd Oblast", "Vologda Oblast", 
    "Voronezh Oblast", "Moscow", "Saint Petersburg", 
    "Sevastopol", "Far Eastern Federal District", 
    "Jewish Autonomous Oblast", "Zabaykalsky Krai", 
    "Ivanovo Oblast", "Irkutsk Oblast", 
    "The Kabardino-Balkar Republic", "Kaliningrad Oblast", 
    "Kaluga Oblast", "Kamchatka Krai", 
    "Karachay-Cherkess Republic", "Kemerovo Oblast", 
    "Kirov Oblast", "Kostroma Oblast", "Krasnodar Krai", 
    "Krasnoyarsk Krai", "Kurgan Oblast", "Kursk Oblast", 
    "Leningrad Oblast", "Lipetsk Oblast", "Magadan District", 
    "Moscow Oblast", "Murmansk Oblast", "Nenets Autonomous Okrug", 
    "Nizhny Novgorod Oblast'", "Novgorod Oblast", 
    "Novosibirsk Oblast", "Omsk Oblast", "Orenburg Oblast", 
    "Oryol Oblast", "Penza Oblast", "Perm Krai", 
    "Volga Federal District", "Primorsky Krai", 
    "Pskov Oblast", "Republic of Adygea", "Altai Republic", 
    "Republic of Bashkortostan", "Republic of Buryatia", 
    "Republic of Dagestan", "Republic of Ingushetia", "Republic of Kalmykia", 
    "Republic of Karelia", "Komi Republic", "Republic of Crimea", 
    "Mari El Republic", "Republic of Mordovia", "Sakha (Yakutia) Republic", 
    "Republic of North Ossetia — Alania", "Republic of Tatarstan", 
    "Tyva Republic", "Republic of Khakassia", "Rostov Oblast", 
    "Ryazan Oblast", "Samara Oblast", "Saratov Oblast", 
    "Sakhalin District", "Sverdlovsk Oblast", 
    "North-Western Federal District", "North Caucasian Federal District", 
    "Siberian Federal District", "Smolensk Oblast", "Stavropol Krai", 
    "Tambov Oblast", "Tver Oblast", "Tomsk District", 
    "Tula Oblast", "Tyumen Oblast", 
    "Tyumen Oblast without Autonomous Okrugs", 
    "Udmurt Republic", "Ulyanovsk Oblast", 
    "Ural Federal District", "Khabarovsk Krai", 
    "Khanty-Mansi Autonomous Okrug — Yugra", 
    "Central Federal District", "Chelyabinsk Oblast", 
    "Chechen Republic", "Chuvash Republic", 
    "Chukotka Autonomous Okrug", "Southern Federal District", 
    "Yamalo-Nenets Autonomous Okrug", "Yaroslavl Oblast"
  )
)
```

##Easy imports (already year-by-region)

###Change in real avg wages from previous year (2018-2023)
```{r}
data_changeinrealavgwages <- read_excel("Macrodata/Raw macrodata/region-year_rosstat_changeinrealavgwages.xlsx", 
                                                       sheet = "с 2018", 
                                                       skip = 2)

data_changeinrealavgwages <- data_changeinrealavgwages %>%
  rename(Region = 1) %>%
  filter(grepl("область|край|республика|Республика|Москва|Петербург|округ", Region))

#Test what doesn't match dictionary
data_changeinrealavgwages %>%
  filter(!(Region %in% region_dict$Russian))

#Prep to standardize
data_changeinrealavgwages <- data_changeinrealavgwages %>%
  mutate(Region = case_when(
    Region == "г. Москва" ~ "г.Москва",
    Region == "в том числе Ненецкий авт.округ" ~ "Ненецкий автономный округ",
    Region == "Архангельская область без авт. округа" ~ "Архангельская область                       без автономного округа",
    Region == "Республика Северная Осетия - Алания" ~ "Республика Северная Осетия-Алания",
    Region == "Ханты-Мансийский авт.округ - Югра" ~ "Ханты-Мансийский автономный                 округ-Югра",
    Region == "Ямало-Ненецкий авт.округ" ~ "Ямало-Ненецкий автономный округ",
    Region == "Тюменская область без авт. округов" ~ "Тюменская область                             без автономных округов",
    Region == "Еврейская авт.область" ~ "Еврейская автономная область",
    Region == "Чукотский авт.округ" ~ "Чукотский автономный округ",
    TRUE ~ Region
  ))

#Check for prep complete
data_changeinrealavgwages %>%
  filter(!(Region %in% region_dict$Russian))

#Check for duplicates
data_changeinrealavgwages %>%
  group_by(Region) %>%
  filter(n() > 1) %>%
  distinct(Region)

#Ready to standardize names
data_changeinrealavgwages <- data_changeinrealavgwages %>%
  left_join(region_dict, by = c("Region" = "Russian")) %>%
  mutate(Region = ifelse(is.na(English), Region, English)) %>%
  select(-English)  # Remove the extra column after replacement

unmatched_regions <- data_changeinrealavgwages %>%
  filter(!(Region %in% region_dict$English))
print(unmatched_regions) #Success!

# Pivot the data to long format
data_changeinrealavgwages <- data_changeinrealavgwages %>%
  pivot_longer(cols = starts_with("20"),  # Select all columns that start with "20"
               names_to = "Year",  # Name of the new column with the years
               values_to = "change_in_real_avg_wages") %>%
  mutate(Year = as.numeric(Year))  # Ensure the Year column is numeric

# View the pivoted data
head(data_changeinrealavgwages)
```

###Poverty rate (2018-2023)
```{r}
data_povertyrate <- read_excel("Macrodata/Raw macrodata/region-year_rosstat_povertyrate.xlsx", 
    sheet = "по РФ и субъектам РФ 2018-2023", 
    skip = 2)

data_povertyrate <- data_povertyrate %>%
  rename(Region = 1) %>%
  filter(grepl("область|край|республика|Республика|Москва|Петербург|округ", Region))

#Test what doesn't match dictionary
data_povertyrate %>%
  filter(!(Region %in% region_dict$Russian))

#Prep to standardize
data_povertyrate <- data_povertyrate %>%
  mutate(Region = case_when(
    Region == "г. Москва" ~ "г.Москва",
    Region == "в том числе: Ненецкий автономный округ" ~ "Ненецкий автономный округ",
    Region == "Архангельская область без авт. округа" ~ "Архангельская область                       без автономного округа",
    Region == "Республика Северная Осетия - Алания" ~ "Республика Северная Осетия-Алания",
    Region == "в том числе: Ханты-Мансийский автономный округ-Югра" ~ "Ханты-Мансийский автономный                 округ-Югра",
    Region == "Ямало-Ненецкий авт.округ" ~ "Ямало-Ненецкий автономный округ",
    Region == "Тюменская область без авт. округов" ~ "Тюменская область                             без автономных округов",
    Region == "Еврейская авт.область" ~ "Еврейская автономная область",
    Region == "Чукотский авт.округ" ~ "Чукотский автономный округ",
    Region == "г. Санкт-Петербург" ~ "г.Санкт-Петербург",
    Region == "Южный федеральный  округ" ~ "Южный федеральный округ",
    TRUE ~ Region
  ))

#Check for prep complete
data_povertyrate %>%
  filter(!(Region %in% region_dict$Russian))

#Check for duplicates
data_povertyrate %>%
  group_by(Region) %>%
  filter(n() > 1) %>%
  distinct(Region)

#Ready to standardize names
data_povertyrate <- data_povertyrate %>%
  left_join(region_dict, by = c("Region" = "Russian")) %>%
  mutate(Region = ifelse(is.na(English), Region, English)) %>%
  select(-English)  # Remove the extra column after replacement

unmatched_regions <- data_povertyrate %>%
  filter(!(Region %in% region_dict$English))
print(unmatched_regions) #Success!

# Pivot the data to long format
data_povertyrate <- data_povertyrate %>%
  pivot_longer(cols = starts_with("20"),  # Select all columns that start with "20"
               names_to = "Year",  # Name of the new column with the years
               values_to = "poverty_rate") %>%
  mutate(Year = as.numeric(Year))  # Ensure the Year column is numeric

# View the pivoted data
head(data_povertyrate)
```

###Year-on-year percentage change in economic output ("index") (2018-2024)
```{r}
data_changeinoutput <- read_excel("Macrodata/Raw macrodata/region-month_rosstat_2022-24_output.xlsx", 
    sheet = "3", skip = 2)

data_changeinoutput <- data_changeinoutput %>%
  rename(Region = 1) %>%
  filter(grepl("область|край|республика|Республика|Москва|Петербург|округ", Region))

#Test what doesn't match dictionary
x <- data_changeinoutput %>%
  filter(!(Region %in% region_dict$Russian))
x$Region

#Prep to standardize
data_changeinoutput <- data_changeinoutput %>%
  mutate(Region = case_when(
    Region == "Тюменская область (без Ханты-Мансийского авт.округа-Югра и Ямало-Ненецкого авт.округа)" ~ "Тюменская область                             без автономных округов",
    Region == "в т.ч. Ненецкий авт. округ" ~ "Ненецкий автономный округ",
    Region == "Архангельская область без Ненецкого авт.округа" ~ "Архангельская область                       без автономного округа",
    Region == "в т.ч. Ханты-Мансийский автономный округ-Югра" ~ "Ханты-Мансийский автономный                 округ-Югра",
    TRUE ~ Region
  ))

#Ready to standardize names
data_changeinoutput <- data_changeinoutput %>%
  left_join(region_dict, by = c("Region" = "Russian")) %>%
  mutate(Region = ifelse(is.na(English), Region, English)) %>%
  select(-English)  # Remove the extra column after replacement

unmatched_regions <- data_changeinoutput %>%
  filter(!(Region %in% region_dict$English))
print(unmatched_regions) #Success!

# Pivot the data to long format
data_changeinoutput[[2]] <- as.numeric(data_changeinoutput[[2]])

data_changeinoutput <- data_changeinoutput %>%
  pivot_longer(cols = starts_with("20"),  # Select all columns that start with "20"
               names_to = "Year",  # Name of the new column with the years
               values_to = "change_in_output") %>%
  mutate(Year = as.numeric(Year))  # Ensure the Year column is numeric

# View the pivoted data
head(data_changeinoutput)
```

###Value of minimum wage for working population (2021-25)
```{r}
data_minwagevalue <- read_excel("Macrodata/Raw macrodata/region-year-value_of_min_wage.xlsx")

data_minwagevalue <- data_minwagevalue %>%
  rename(Region = 1) %>%
  filter(grepl("область|край|республика|Республика|Москва|Петербург|округ", Region))

#Test what doesn't match dictionary
x <- data_minwagevalue %>%
  filter(!(Region %in% region_dict$Russian))
x$Region

#Prep to standardize
data_minwagevalue <- data_minwagevalue %>%
  mutate(Region = case_when(
    Region == "г. Москва" ~ "г.Москва",
    Region == "г. Санкт-Петербург" ~ "г.Санкт-Петербург",
    Region == "Архангельская область1)" ~ "Архангельская область",
    Region == "Hенецкий авт.округ" ~ "Ненецкий автономный округ",
    Region == "Архангельская область без авт. округа" ~ "Архангельская область                       без автономного округа",
    Region == "Республика Северная Осетия- Алания" ~ "Республика Северная Осетия-Алания",
    Region == "Hижегородская область" ~ "Нижегородская область",
    Region == "Тюменская область1)" ~ "Тюменская область",
    Region == "Ханты-Мансийский авт.округ-Югра" ~ "Ханты-Мансийский автономный                 округ-Югра",
    Region == "Ямало-Hенецкий авт.округ" ~ "Ямало-Ненецкий автономный округ",
    Region == "Тюменская область без авт. округов" ~ "Тюменская область                             без автономных округов",
    Region == "Чукотский авт.округ" ~ "Чукотский автономный округ",
    TRUE ~ Region
  ))

#Ready to standardize names
data_minwagevalue <- data_minwagevalue %>%
  left_join(region_dict, by = c("Region" = "Russian")) %>%
  mutate(Region = ifelse(is.na(English), Region, English)) %>%
  select(-English)  # Remove the extra column after replacement

unmatched_regions <- data_minwagevalue %>%
  filter(!(Region %in% region_dict$English))
print(unmatched_regions) #Success!

# Pivot the data to long format
data_minwagevalue[, 2:ncol(data_minwagevalue)] <- lapply(data_minwagevalue[, 2:ncol(data_minwagevalue)], as.numeric)

data_minwagevalue <- data_minwagevalue %>%
  pivot_longer(cols = starts_with("20"),  # Select all columns that start with "20"
               names_to = "Year",  # Name of the new column with the years
               values_to = "minwagevalue") %>%
  mutate(Year = as.numeric(Year))  # Ensure the Year column is numeric

#Arkangelskaya oblast doesn't have any data for some reason. Carry over the "without autonomous okrug" values
data_minwagevalue <- data_minwagevalue %>%
  group_by(Year) %>%
  mutate(minwagevalue = if_else(Region == "Arkhangel'skaya Oblast'" & is.na(minwagevalue), 
                                minwagevalue[Region == "Arkhangelskaya Oblast without Autonomous Okrug"], 
                                minwagevalue)) %>%
  mutate(minwagevalue = if_else(Region == "Tyumen Oblast" & is.na(minwagevalue), 
                                minwagevalue[Region == "Tyumen Oblast without Autonomous Okrugs"], 
                                minwagevalue)) %>%
  ungroup()
```

###Handle missing data

###Mediazona casualty data (2022-24)
```{r}
data_mediazona <- read_excel("Macrodata/Mediazona.xlsx", 
    sheet = "Region-by-year")

data_mediazona <- data_mediazona %>%
  pivot_longer(cols = -Period,  # Select all columns that start with "20"
               names_to = "Region",  # Name of the new column with the years
               values_to = "mediazona") %>%
  mutate(Year = as.numeric(Period)) %>% # Ensure the Year column is numeric
  select(-Period)

#Test what doesn't match dictionary
x <- data_changeinoutput %>%
  filter(!(Region %in% region_dict$English))
x$Region
#perfect (by construction√)

# View the pivoted data
head(data_mediazona)
```


##Industrial output indices (2015-2014, monthly)

###Load the data in (loop over 5 sheets, merge years/months, clean)
```{r}
# Define sheet numbers to loop over
sheets_to_read <- c("1", "4", "7", "10", "13")

# Create an empty list to store cleaned data
cleaned_data_list <- list()

# Loop over each sheet
for (sheet_num in sheets_to_read) {
  
  # Step 1: Import data
  raw_data <- read_excel("Macrodata/Raw macrodata/region-month_rosstat_industrialoutput.xlsx", 
                         sheet = sheet_num, skip = 1, col_names = FALSE)  # Read without headers
  
  # Step 2: Extract relevant parts
  years <- as.character(unlist(raw_data[3, ]))  # Extract year row
  months <- as.character(unlist(raw_data[4, ])) # Extract month row
  
  # **NEW**: Remove superscript footnotes and extra characters
  months <- str_replace_all(months, "[^а-яА-Я]", "")  # Keep only Russian letters
  
  # Step 3: Propagate year values
  year_filled <- na.locf(years, na.rm = FALSE) # Fill down nearest year
  year_filled <- make.names(year_filled, unique = TRUE)  # Add .a, .b, etc.
  
  # Step 4: Create final column names
  column_names <- paste(year_filled, months, sep = "_")
  
  # Step 5: Assign new column names and clean up data
  cleaned_data <- raw_data[-c(1:4), ]  # Remove first 4 rows (headers)
  colnames(cleaned_data) <- c("Region", column_names[-1])  # Keep "Region" as the first column
  
  # Step 6: Extract year and month separately
  cleaned_columns <- str_replace_all(colnames(cleaned_data), "X|\\.год|\\.\\d+", "")  # Remove extra text
  month_year_split <- str_match(cleaned_columns, "(\\d{4})_(.*)")  # Extract Year and Month parts
  years <- month_year_split[, 2]  # Year column
  months_russian <- month_year_split[, 3]  # Month column
  
  # **NEW**: Clean up month names again before translation
  months_russian <- str_replace_all(months_russian, "[^а-яА-Я]", "")  # Keep only Russian letters
  
  # Step 7: Translate months
  month_translation <- c("январь" = "January", "февраль" = "February", "март" = "March",
                         "апрель" = "April", "май" = "May", "июнь" = "June",
                         "июль" = "July", "август" = "August", "сентябрь" = "September",
                         "октябрь" = "October", "ноябрь" = "November", "декабрь" = "December")
  
  months_english <- recode(months_russian, !!!month_translation)  # Convert Russian months to English
  
  # Step 8: Create new formatted names
  new_column_names <- ifelse(!is.na(years) & !is.na(months_english), paste(months_english, years), "Region")
  
  # Step 9: Assign cleaned column names
  colnames(cleaned_data) <- new_column_names
  
  # Step 10: Store the cleaned data in a list
  cleaned_data_list[[paste0("Sheet_", sheet_num)]] <- cleaned_data
}

# Display the first few rows of each cleaned dataset
lapply(cleaned_data_list, head)
```

###Clean the region names
```{r}
# Loop through each data frame in cleaned_data_list
for (sheet_name in names(cleaned_data_list)) {
  cleaned_data_list[[sheet_name]] <- cleaned_data_list[[sheet_name]] %>%
    rename(Region = 1) %>%
    filter(grepl("область|край|республика|Республика|Москва|Петербург|округ", Region))
  
  # Test what doesn't match dictionary
  x <- cleaned_data_list[[sheet_name]] %>%
    filter(!(Region %in% region_dict$Russian))
  #print(x$Region)
  
  #Prep to standardize
  cleaned_data_list[[sheet_name]] <- cleaned_data_list[[sheet_name]] %>%
    mutate(Region = case_when(
      Region == "г. Москва" ~ "г.Москва",
      Region == "в том числе Ненецкий авт.округ" ~ "Ненецкий автономный округ",
      Region == "в том числе               Ненецкий авт.округ" ~ "Ненецкий автономный округ",
      Region == "в том числе               Ненецкий авт. округ" ~ "Ненецкий автономный округ",
      Region == "в том числе               \nНенецкий авт.округ" ~ "Ненецкий автономный округ",
      Region == "Архангельская область без авт. округа" ~ "Архангельская область                       без автономного округа",
      Region == "г. Санкт-Петербург" ~ "г.Санкт-Петербург",
      Region == "Республика Северная  Осетия - Алания" ~ "Республика Северная Осетия-Алания",
      Region == "Приволжский               федеральный округ" ~ "Приволжский федеральный округ",
      Region == "Уральский  федеральный округ" ~ "Уральский федеральный округ",
      Region == "Уральский             федеральный округ" ~ "Уральский федеральный округ",
      Region == "в том числе: Ханты-Мансийский авт. округ - Югра" ~ "Ханты-Мансийский автономный                 округ-Югра",
      Region == "в том числе:                     Ханты-Мансийский       авт. округ - Югра" ~ "Ханты-Мансийский автономный                 округ-Югра",
      Region == "в том числе:                     \nХанты-Мансийский авт. округ - Югра" ~ "Ханты-Мансийский автономный                 округ-Югра",
      Region == "Ямало-Ненецкий авт. округ" ~ "Ямало-Ненецкий автономный округ",
      Region == "Ямало-Ненецкий             авт. округ" ~ "Ямало-Ненецкий автономный округ",
      Region == "Тюменская область без авт. округов" ~ "Тюменская область                             без автономных округов",
      Region == "Еврейская авт.область" ~ "Еврейская автономная область",
      Region == "Еврейская авт. область" ~ "Еврейская автономная область",
      Region == "Чукотский авт.округ" ~ "Чукотский автономный округ",
      Region == "Чукотский авт. округ" ~ "Чукотский автономный округ",
      Region == "Южный                   федеральный округ" ~ "Change",
      Region == "Сибирский           федеральный округ" ~ "Сибирский федеральный округ",
      TRUE ~ Region
    ))
  
  #Ready to standardize names
  cleaned_data_list[[sheet_name]] <- cleaned_data_list[[sheet_name]] %>%
    left_join(region_dict, by = c("Region" = "Russian")) %>%
    mutate(Region = ifelse(is.na(English), Region, English)) %>%
    select(-English)  # Remove the extra column after replacement
  
  unmatched_regions <- cleaned_data_list[[sheet_name]] %>%
    filter(!(Region %in% region_dict$English))
  print(unmatched_regions$Region) #Success!
  
}
```

###Rename the datasets
```{r}
# Rename the list
data_changeinindustrialproduction <- cleaned_data_list

# Define new names for each sheet
sheet_names <- c("total", "mineral_extraction", "manufacturing",
                 "electricity_utilities", "water_utilities")

# Rename list elements and modify each data frame
names(data_changeinindustrialproduction) <- sheet_names
```

###Convert to the proper format and merge
```{r}
final_data <- list()

# Loop through each sheet in the cleaned data list
for (sheet_name in names(cleaned_data_list)) {
  
  # Step 1: Convert the dataset from wide to long format
  data_long <- cleaned_data_list[[sheet_name]] %>%
    pivot_longer(
      cols = starts_with("January 2015"):ends_with("December 2024"), 
      names_to = "Month_Year", 
      values_to = sheet_name  # Use the sheet name as the column name
    )
  
  # Step 2: Convert the "Month_Year" column to Date format (assuming it's in the "Month Year" format)
  data_long <- data_long %>%
    mutate(Month_Year = my(Month_Year))
  
  # Step 3: Add the data to the final_data list
  final_data[[sheet_name]] <- data_long
}

# Step 4: Merge all the data into a single dataset
# We'll perform a full join on the common columns "Region" and "Month_Year"
final_data_merged <- reduce(final_data, full_join, by = c("Region", "Month_Year")) %>%
  mutate(changeinindustrialproduction_all = Sheet_1,
         changeinindustrialproduction_mineral_extraction = Sheet_4,
         changeinindustrialproduction_manufacturing = Sheet_7,
         changeinindustrialproduction_electricity_utilities = Sheet_10,
         changeinindustrialproduction_water_utilities = Sheet_13
         ) %>%
  select(-Sheet_1, -Sheet_4, -Sheet_7, -Sheet_10, -Sheet_13)
```

###Annualize
```{r}
#Export a clean monthly version for potential later use
write_csv(final_data_merged, "Macrodata/Cleaned macrodata (intermediate)/monthly_changeinindustrialproduction.csv")

#Aggregate to annual level by taking the december value
#unit is increase since the same month in the previous year, so this is an annual increase.
data_changeinindustrialproduction <- final_data_merged %>%
  # Extract the year and month from the Month_Year column
  mutate(Year = lubridate::year(Month_Year), Month = lubridate::month(Month_Year)) %>%
  # Filter for December (Month == 12)
  filter(Month == 12) %>%
  # Select the necessary columns
  select(Region, Year, starts_with("changeinindustrialproduction")) %>%
  # Ungroup to remove grouping
  ungroup()

# Replace "-" with 100 in the column 'changeinindustrialproduction_mineral_extraction'
data_changeinindustrialproduction <- data_changeinindustrialproduction %>%
  mutate(changeinindustrialproduction_mineral_extraction = as.numeric(changeinindustrialproduction_mineral_extraction),
         changeinindustrialproduction_mineral_extraction = 
           if_else(is.na(changeinindustrialproduction_mineral_extraction), 
                   100, 
                   changeinindustrialproduction_mineral_extraction))
#represents moscow, which doesn't have mineral extraction and is an NA. 

head(data_changeinindustrialproduction)
```

##Population
```{r}
library(readxl)
library(dplyr)
library(stringr)

# Define file path
file_path <- "Macrodata/Raw macrodata/region-year_Daria_population.xlsx"

# Get sheet names (years)
sheet_names <- excel_sheets(file_path)

# Function to read and process each sheet
read_population_data <- function(sheet) {
  data <- read_excel(file_path, sheet = sheet, col_names = TRUE) %>%
    filter(str_detect(Region, "область|край|республика|Республика|Москва|Петербург|автономный")) %>%
    mutate(Year = as.numeric(sheet)) # Convert sheet name to Year
  
  return(data)
}

# Import and combine all sheets
population_data <- bind_rows(lapply(sheet_names, read_population_data))

# Reshape the data so that each row is a region-year observation
population_data_clean <- population_data %>%
  select(-`...1`) %>%  # Remove unnecessary column
  relocate(Year, .after = Region)  # Ensure Year comes right after Region

# View the first few rows
head(population_data_clean)
```

###Translate the regions
```{r}
population_data_clean %>%
  filter(!(Region %in% region_dict$Russian)) %>%
  distinct(Region)

# Standardize region names
population_data_clean <- population_data_clean %>%
  mutate(Region = case_when(
    Region == "г. Москва" ~ "г.Москва",
    Region == "г. Санкт-Петербург - город федерального значения" ~ "г.Санкт-Петербург",
    Region == "Архангельская область без авт. округа" ~ "Архангельская область",
    Region == "Архангельская область, включая Ненецкий автономный округ" ~ "Архангельская область",
    Region == "Тюменская область включая автономные округа" ~ "Тюменская область",
    Region == "Тюменская область без автономных округов" ~ "Тюменская область                             без автономных округов",
    Region == "Ханты-Мансийский автономный округ - Югра" ~ "Ханты-Мансийский автономный                 округ-Югра",
    Region == "Город федерального значения - г. Санкт-Петербург" ~ "г.Санкт-Петербург",
    Region == "Город федерального значения - г. Москва" ~ "г.Москва",
    Region == "г. Москва - город федерального значения" ~ "г.Москва",
    Region == "Архангельская область без Ненецкого автономного округа" ~ "Архангельская область                       без автономного округа",
    Region == "Тюменская область (кроме Ханты-Мансийского автономного округа - Югры и Ямало-Ненецкого автономного округа)" ~ "Тюменская область                             без автономных округов",
    Region == "Кемеровская область - Кузбасс" ~ "Кемеровская область",
    Region == "г Москва - город федерального значения" ~ "г.Москва",
    Region == "г Санкт-Петербург - город федерального значения" ~ "г.Санкт-Петербург",
    TRUE ~ Region
  ))

region_dict %>%
  filter(!(Russian %in% population_data_clean$Region))
#great, we don't need the federal okrug data, and Sevastopol is Ukrainian 

# Check for duplicates
population_data_clean %>%
  group_by(Region, Year) %>%
  filter(n()>1) 
#same, just keep the first occurrence 
population_data_clean <- population_data_clean %>%
  distinct(Region, Year, .keep_all = TRUE)

# Standardize names using the dictionary
population_data_clean <- population_data_clean %>%
  left_join(region_dict, by = c("Region" = "Russian")) %>%
  mutate(Region = ifelse(is.na(English), Region, English)) %>%
  select(-English)  # Remove the extra column after replacement

# Check if any regions are still unmatched
unmatched_regions <- population_data_clean %>%
  filter(!(Region %in% region_dict$English))
print(unmatched_regions)  # Should return an empty tibble if successful

# View the cleaned data
head(population_data_clean)

data_population <- population_data_clean
```

###Handle NAs
```{r}
# Replace NA with 0 in the column 'rural_pop'
data_population <- data_population %>%
  mutate(rural_pop = 
           if_else(is.na(rural_pop), 
                   0, 
                   rural_pop))
#represents cites, which don't have rural population and would be an NA. 
```

##Average wages (2012-2024, monthly)
```{r}
library(readxl)
library(tidyr)
library(dplyr)
library(lubridate)

# Step 1: Read the data
data_avgwages <- read_excel("/Users/michelle/Documents/thesis/Macrodata/Raw macrodata/region-month_rosstat_avgwages.xlsx", sheet = 1)

# Step 2: Extract the month names from the second row and the years from the column names
months <- data_avgwages[1, -1]  # Remove the first column (Region column)
years <- colnames(data_avgwages)[-1]  # Remove the first column (Region column)
years <- sub("\\.\\.\\..*", "", years)

# Step 3: Clean up the data
data_clean <- data_avgwages[-1, ]  # Remove the first two rows
colnames(data_clean) <- c("Region", paste(months, years, sep = "-"))  # Assign month names as column names
data_clean <- pivot_longer(data_clean, cols = -Region, names_to = "Month", values_to = "avg_nominal_gross_wages")

data_clean <- data_clean %>%
  mutate(Date = dmy(paste0("01-", Month))) %>%  # Convert to date format
  select(Region, Date, avg_nominal_gross_wages)
```

###Translate the regions
```{r}
data_clean %>%
  filter(!(Region %in% region_dict$Russian)) %>%
  distinct(Region)

# Standardize region names
data_clean <- data_clean %>%
  mutate(Region = case_when(
    Region == "Архангельская область без авт. округа." ~ "Архангельская область                       без автономного округа",
    Region == "Тюменская область без авт. округов" ~ "Тюменская область                             без автономных округов",
    Region == "Республика Северная  Осетия - Алания" ~ "Республика Северная Осетия-Алания",
    Region == "Ямало-Ненецкий авт. округ" ~ "Ямало-Ненецкий автономный округ",
    Region == "Еврейская авт.область" ~ "Еврейская автономная область",
    Region == "Чукотский авт.округ" ~ "Чукотский автономный округ",
    Region == "Ханты-Мансийский  авт. округ - Югра" ~ "Ханты-Мансийский автономный                 округ-Югра",
    Region == "Российская Федерация" ~ "Russian Federation",
    TRUE ~ Region
  ))

region_dict %>%
  filter(!(Russian %in% data_clean$Region))

# Check for duplicates
data_clean %>%
  group_by(Region, Date) %>%
  filter(n()>1) 

# Standardize names using the dictionary
data_clean <- data_clean %>%
  left_join(region_dict, by = c("Region" = "Russian")) %>%
  mutate(Region = ifelse(is.na(English), Region, English)) %>%
  select(-English) # Remove the extra column after replacement

# Check if any regions are still unmatched
unmatched_regions <- data_clean %>%
  filter(!(Region %in% region_dict$English)) %>%
  distinct(Region)
print(unmatched_regions)  # Should return an empty tibble if successful

# View the cleaned data
head(data_clean)

data_avgnominalgrosswages <- data_clean
```

###Annualize
```{r}
#Export a clean monthly version for potential later use
write_csv(data_avgnominalgrosswages, "Macrodata/Cleaned macrodata (intermediate)/monthly_avgnominalgrosswages.csv")

data_avgnominalgrosswages <- data_avgnominalgrosswages %>%
  mutate(avg_nominal_gross_wages = as.numeric(avg_nominal_gross_wages)) %>%
  mutate(Year = year(Date)) %>%  # Extract year from Date column
  group_by(Region, Year) %>%
  summarise(avg_nominal_gross_wages = mean(avg_nominal_gross_wages, na.rm = TRUE), .groups = "drop")  # Annualize by averaging

head(data_avgnominalgrosswages)
```


#Merge data to existing flows/yandex dataframe

##All sets imported and standardized (*=aggregated): 
data_povertyrate (region/year, 2018-23)
data_changeinoutput (region/year, 2018-23)
data_changeinrealavgwages (region/year, 2018-23)
data_minwagevalue (region/year, 2021-25)
data_mediazona (region/year, 2022-24)
data_changeinindustrialproduction (region/year*, 2015-24)
data_population (region/year*, 2013-24)
data_avgnominalgrosswages (region/year*, 2013-24)

```{r}
merged_data <- read_csv("merged_data.csv")

merged_data %>%
  filter(!(Region %in% region_dict$English)) %>%
  distinct(Region) #ready to go

merged_data <- merged_data %>%
  left_join(data_povertyrate, by = c("Region", "Year")) %>%
  left_join(data_changeinoutput, by = c("Region", "Year")) %>%
  left_join(data_changeinrealavgwages, by = c("Region", "Year")) %>%
  left_join(data_minwagevalue, by = c("Region", "Year")) %>%
  left_join(data_mediazona, by = c("Region", "Year")) %>%
  left_join(data_changeinindustrialproduction, by = c("Region", "Year")) %>%
  left_join(data_population, by = c("Region", "Year")) %>%
  left_join(data_avgnominalgrosswages, by = c("Region", "Year"))

#handle NAs for mediazona
merged_data$mediazona[is.na(merged_data$mediazona)] <- 0

merged_data <- merged_data %>%
  mutate(log_minwagevalue = log1p(minwagevalue),
         log_mediazona = log1p(mediazona),
         log_total_pop = log1p(total_pop),
         log_city_pop = log1p(city_pop),
         log_rural_pop = log1p(rural_pop),
         log_avg_nominal_gross_wages = log1p(avg_nominal_gross_wages))

merged_data <- merged_data %>%
  mutate(across(.cols = 172:ncol(merged_data),  # Select columns starting from column 172
                .fns = list(lag = ~lag(.)),      # Apply lag function
                .names = "lag_{.col}"))          # Name the new column as lag_[variablename]

unique_years <- unique(merged_data$Year)
print(unique_years)


```

##Export finally!
```{r}
write.csv(merged_data, "region-year-data.csv", row.names = FALSE)
```

